### 2013-06-18 conclusion.tex
first commit

\section{Conclusion and future work}
\label{sec:concl-future-work}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: