### 2013-07-01 background.tex
Included all changes by Prof. Salcic

\section{Background}
\label{sec:background}

\subsection{Brief introduction to SystemJ syntax, semantics and model of
  computation}
\label{sec:brief-intr-syst}

\begin{table}[tb]
\centering
\caption{SystemJ kernel statements and their meaning}
\begin{minipage}{8cm}
  \begin{scriptsize}
   \begin{tabular}{|c|p{80pt}|}
     \hline                                                                                     
     \textbf{Kernel Statements} & \textbf{Meaning}\\                                            
     \hline                                                                                     
     \hline                                                                                     
     [\textbf{\texttt{input}}] [\textbf{\texttt{output}}]
     [\textbf{\texttt{type}}] \textbf{\texttt{signal}} S & declare signal S\\                                      
     \hline                                                                                     
     \textbf{\texttt{emit}} S [(value)] & broadcast signal S\\                                                    
     \hline                                                                                     
     \textbf{\texttt{present}} (S) \{p\} else \{q\}& do p if S is present, else do q\\                            
     \hline                                                                                     
     \textbf{\texttt{abort}} (S) \{p\} & preempt program p if S is present\\                                      
     \hline                                                                                     
     \textbf{\texttt{suspend}} (S)\{p\} & suspend p if S is present\\                                             
     \hline                                                                                     
     \textbf{\texttt{trap}} (T)\{p\ldots \textbf{\texttt{exit}} T\ldots\} & preempt p if exit is executed\\                         
     \hline                                                                                     
     p\textbf{\texttt{$||$}}q & run p and q in lock-step\\                                                        
     \hline                                                                                     
     p$><$q & run p and q asynchronously\\                                                      
     \hline                                                                                     
     \textbf{\texttt{send}} C([value]) & send a value through C, blocking
     send\\                                                 
     \hline                                                                                     
     \textbf{\texttt{receive}} C() & receive a value through C, blocking
     receive\\
     \hline                                                                                     
     \textbf{\texttt{pause}} & finish a tick and communicate
     with environment\\
     \hline                                                                                     
   \end{tabular}
  \end{scriptsize}
   % \footnotetext[1]{\scriptsize Uppercase}
 \end{minipage}
 \label{tab:1}
\end{table}

Table~\ref{tab:1} presents the SystemJ kernel statements used to program
the control flow. The data flow is programmed in Java. A number of
derived statements also exist (e.g., \texttt{sustain}, \texttt{await},
etc) that make programming reactive systems easier. The \textit{Model of
  Computation} (MoC) of a simple SystemJ program is shown in
Figure~\ref{fig:2}.

SystemJ consists of entities called clock-domains (CD)s, each running at
their individual pace, interacting with the environment via signals and
with each other via channels and CSP~\cite{choa85} style
rendezvous. Each CD has its own notion of logical time. CDs adheres to
the \textit{perfect synchrony} hypothesis, i.e., all statements execute
instantaneously in zero time. Only the \texttt{pause} statement consumes
time, just like in Esterel~\cite{gber931}. Consider the SystemJ program
in Figure~\ref{fig:2a}, it is waiting for an input signal producing
logical ticks (see Figure~\ref{fig:2c}). Once \texttt{A} is received at
tick 1 or 4, output \texttt{O} is emitted to the environment
instantaneously.

\subsection{Mapping logical time to physical time}
\label{sec:mapping-logical-time}

\begin{figure}[t!]
% \centering
\begin{SubFloat}{\label{fig:2a}Simple SystemJ program}%...verbatim subfigure
\begin{minipage}[b]{0.3\linewidth}% a minipage to control the width...
  \begin{scriptsize}
\begin{verbatim}
{
int counter = 0;
while(true){
 abort(A){
  while(true) pause; 
  emit O;
  if(counter != 0) { // do java computation leading to WCRT}
  ++counter;
  } // abort end
 } // loop end
} // SystemJ CD end
\end{verbatim}%
  \end{scriptsize}
\end{minipage}%
\end{SubFloat}
\hspace{1.5cm}%
% \begin{SubFloat}[Black box]{\label{fig:2b}Logical time in SystemJ}%
% \includegraphics[scale=0.4]{moc}
% \end{SubFloat}%
% \vspace{-10pt}

\begin{SubFloat}{\label{fig:2c}Mapping to physical time}%
% \includegraphics[scale=0.4]{phy.pdf_t}
\scalebox{0.68}{\input{phy.pdf_t}}
\end{SubFloat}%
\caption{Simple SystemJ example and corresponding MoC}
\label{fig:2}
\end{figure}

The perfect synchrony hypothesis is ideal for programming the
synchronous sub-set of SystemJ. Unfortunately, every \textit{micro-step}
requires real-time to compute. Formally, execution of every micro-step
in the logical zero time, requires $\delta$ physical time. Time for a
logical tick can thus be summarized as $\Delta = \mathcal{F} (\delta)$,
where $\mathcal{F}$ is some function dependent upon the schedule. We
call $\Delta$ the reaction time. Obviously, depending upon the amount of
computation required and schedule, $\Delta$ can vary, hence, in order to
satisfy the implicit restriction imposed by the synchrony hypothesis --
no input event \textit{from the environment} can be missed -- one needs
to calculate the \textit{Worst Case Reaction Time} (WCRT) and the
resultant WCRT needs to be smaller than the time between any two
consecutive events on any input, else the synchrony hypothesis is
violated. Techniques for calculating the WCRT of synchronous
programs~\cite{boldt07} exist and this analysis is not the focus of this
paper. The opposite of the WCRT is the \textit{Best Case Reaction Time}
(BCRT), we denote the WCRT and BCRT in Figure~\ref{fig:2c} using
\texttt{W} and \texttt{B}, respectively. Figure~\ref{fig:2c} shows a
continuously running physical clock (analogous to a clock in digital
hardware), the numerical annotations on the rising edge of the clock
mark the logical ticks from Figure~\ref{fig:2c}. The subscripts
\texttt{s} and \texttt{e} for the numbers show the start and end of the
logical ticks, respectively. The end of a logical tick and the start of
the next logical tick happen together. Figure~\ref{fig:2c} shows the
mapping of the logical time to the physical time. For example, the
logical tick \texttt{1} starts at the first rising edge and completes at
the second rising edge of the physical clock. Where as logical tick
\texttt{4} starts at the $4^{th}$ rising edge, but finishes at the
$6^{th}$ rising edge of the physical clock. This difference can be
accounted for by the java computations required when signal \texttt{A}
is input by the environment (see Figure~\ref{fig:2a}). This elasticity
is inherent (and elegant) to both Esterel and SystemJ.

% We now present a lemma with proof sketch (due to lack of space) that we
% require for the rest of the paper and is special to SystemJ, being a
% GALS language.

% \begin{lemma}
%   WCRT and BCRT analysis are invariant to values of conditional
%   expressions.
% \end{lemma}

% \begin{proof}
%   The physical time ($\delta$) taken by a conditional expression is at
%   best affected by the type rather than the value of the
%   conditional. Example, comparing the value of 2 floats might take
%   longer than comparing the value of 2 integer types on some given
%   platform. But, the time taken by the comparison instruction is
%   invariant to the value itself.
% \end{proof}

% \begin{lemma}
%   WCRT and BCRT analysis are invariant to channel communication.
% \end{lemma}
% \begin{proof}
%   LATER ....
% \end{proof}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: