### 2013-07-01 related_work.tex

\section{Related work}
\label{sec:related-work}

A small number of studies have been performed when it comes to
integrating real-time directly as a programming construct in GALS (and
its subset synchronous) reactive languages. The most prominent works in
this area are by Shyamsundar~\cite{rsh94} and Bourke et
al.~\cite{Bourke2009a}. Shyamsundar incorporates real-time using
external timers in \textit{Communicating Reactive Processes} (CRP),
which like SystemJ is an extension of synchronous language Esterel to
asynchorny. But, as mentioned in Section~\ref{sec:inter-preempt-delays},
external timers do not interact well with preemption constructs in these
languages. Bourke et al. introduce real-time as first class constructs
in the Esterel language. They like us provide real-time \texttt{delay}
as first class Esterel programming constructs, and translate them into
Esterel kernel constructs. But, unlike us they do not translate delays
into \texttt{pause} constructs directly. Instead, logical ticks are
generated by using abstract notion of \texttt{event} and \texttt{sample}
platform dependent timers. This notion makes it a complex and inflexible
solution since the number of platform timers with certain resolutions
need to be determined and present on the system for the solution to be
realizable. Moreover, non-deterministic real-time delays and integration
with non-maximal parallelism is not studied at all. Our solution does
not require external timers (like~\cite{rsh94}) or logical notion of
timers (like~\cite{Bourke2009a}). Since in our solution we calculate the
delay in logical ticks \textit{after} allocation and scheduling, it
gives the compiler developers the chance to optimize for many different
criteria such as computation time, energy, power, etc, without worrying
about or violating the real-time delay specification. Other works such
as Quartz~\cite{glog02} also incorporate delay statements in Esterel
using \texttt{pause} constructs, but again, non-maximal parallelism and
non-deterministic delays are not studied. Moreover, Quartz is targeted
at studying timing properties using model checkers rather than
implementation, like us.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: