### 2013-11-17 conclusion.tex

\section{Conclusion}
\label{sec:concl-future-work}

In this paper we have described a novel way to introduce real-time in
\textit{Globally Asynchronous Locally Synchronous} (GALS) languages, in
particular SystemJ, and their subset the synchronous languages like
Esterel. The fundamental idea is to convert real-time delay into logical
delays (\texttt{pause} construct) that interact well with rest of the
constructs in these languages, especially, preemption and channel
communication. We do \textit{not} need or use external timers to
introduce these real-time delays, instead we depend upon an underlying
platform that is timing analyzable, thereby resolving the problems of
interaction between external timers and preemption. Moreover, we also do
not require specific timer resolutions, thereby making our solution
elegant (in the spirit of GALS/synchronous programming) and flexible. We
rewrite the delay construct into reactive kernel constructs, calculate
the \textit{Worst} and \textit{Best} case reaction times and then
determine \textit{delay} \texttt{d}, number of logical ticks, which is
invariant to the \textit{elasticity} of logical time. To our knowledge
we are the first to introduce precise and imprecise delays in such
languages and show their usefulness in designing real-world systems.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: