### 2013-11-17 related_work.tex

\section{Related work}
\label{sec:related-work}

A number of studies have been performed when it comes to integrating
real-time directly as a programming construct in GALS (and its subset
synchronous) reactive languages. The most prominent works in this area
are by Shyamsundar~\cite{rsh94} and Bourke et
al.~\cite{Bourke2009a}. Shyamsundar incorporates real-time using
external timers in \textit{Communicating Reactive Processes} (CRP),
which like SystemJ is an extension of synchronous language Esterel to
asynchorny. But, as mentioned in Section~\ref{sec:inter-preempt-delays},
external timers do not interact well with preemption constructs in these
languages. Bourke et al. introduce real-time as first class constructs
in the Esterel language. They like us provide real-time \texttt{delay}
as first class Esterel programming constructs, and translate them into
Esterel kernel constructs. But, unlike us they do not translate delays
into \texttt{pause} constructs directly. Instead, logical ticks are
generated by using abstract notion of \texttt{event} and \texttt{sample}
platform dependent timers. This notion makes it a complex and inflexible
solution since the number of platform timers with certain resolutions
need to be determined and present on the system for the solution to be
realizable. Moreover, non-deterministic real-time delays and integration
with non-maximal parallelism is not studied at all. Our solution does
not require external timers (like~\cite{rsh94}) or logical notion of
timers (like~\cite{Bourke2009a}). Since in our solution we calculate the
delay in logical ticks \textit{after} allocation and scheduling, it
gives the compiler developers the chance to optimize for many different
criteria such as computation time, energy, power, etc, without worrying
about or violating the real-time delay specification.

Other works such as Quartz~\cite{glog02} and Bertin
et.al.~\cite{Bertin:2000:TVR:1947412.1947439}, also incorporate delay
statements in Esterel using \texttt{pause} and special \textit{pragmas}
(annotations), respectively.  The Quartz solution is targeted at
studying timing properties using model checkers rather than
implementation, like us. The solution introduced by Bertin
et.al.~\cite{Bertin:2000:TVR:1947412.1947439} translates real-time
pragmas into Timed automata~\cite{alur94} for model-checking real-time
deadlines. Our solution on the other hand removes all continuous time
elements and produces single discrete time model thereby making the
solution amenable to existing functional and real-time verification
tools developed for reactive languages and more in the spirit of logical
time. Finally, a major difference between Bertin et.al and our approach
is that their approach requires programmers to annotate the time taken
by 'C' code fragments manually, this is extremely error prone, we use
static low-level program analysis to find out such times automatically.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: