### 2014-01-27 conclusion.tex

\section{Conclusion}
\label{sec:concl-future-work}

In this paper we have described a novel way to introduce real-time in
\textit{Globally Asynchronous Locally Synchronous} (GALS) languages, in
particular SystemJ, and their subset the synchronous languages like
Esterel. The fundamental idea is to convert real-time wait constructs
into logical waits (\texttt{pause} construct) that interact well with
rest of the constructs in these languages, especially, preemption and
channel communication. We do \textit{not} need or use external timers to
introduce these real-time wait constructs, instead we depend upon an
underlying platform that is timing analyzable, thereby resolving the
problems of interaction between external timers and
preemption. Moreover, we also do not require specific timer resolutions,
thereby making our solution elegant (in the spirit of GALS/synchronous
programming) and flexible. We rewrite the wait constructs into reactive
kernel constructs, calculate the \textit{Worst} and \textit{Best} case
reaction times and then determine wait \texttt{d} in number of logical
ticks, which is invariant to the \textit{elasticity} of logical time. To
our knowledge we are the first to introduce exact and inexact wait
constructs in such languages and show their usefulness in designing
real-world systems.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: