### 2014-01-27 motivation.tex

\subsection{Motivation via examples of real-time systems programmed in a
reactive language -- SystemJ}
\label{sec:motivating-example}

% system{
%  interface{
%   input signal TOUCH;
%   output signal GREEN_LIGHT;
%   output char channel S;
%   input char channel R;
%  }

Real-time operating systems~\cite{barry2009using} usually provide two
types of mechanisms for introducing real-time in the developed program:
(1) the ability to perform a timeout and (2) the ability to run tasks
periodically with some real-time period $T$. We would like to provide
similar mechanisms in reactive languages for real-time program
development. In fact, we provide \textit{exact} and \textit{non exact}
real-time control mechanisms -- a generalization of the two real-time
mechanisms introduced in real-time operating systems.

\subsubsection{Programming using the non exact real-time control
  construct the \textrm{\texttt{wait\_inbetween}} statement}
\label{sec:progr-using-non}

\begin{figure}[t!]
	\vspace{-10pt}
        \begin{SubFloat}{\label{delaycode:a}A human responsiveness
            system (HRTCS)}
        \begin{lstlisting}[style=sysj,morekeywords={sustain,send,receive,abort,await,emit,present,trap,pause,exit,wait_inbetween,suspend}]
{ // Clock-domain 1
while(true) {
  trap(TRAP){
   // Abort if user touches the screen
   abort(TOUCH){
    {sustain GREEN_LIGHT;} //reaction R1
    || // synchronous parallel operator
    {
     //exit after any where from 50.3 to 200.3 ms.
     wait_inbetween (50.3 .. 200.3 ms); 
     exit(TRAP); 
    } // reaction R2
   } do {send S(1);} // Send 1 if touched 
  } do {send S(0);} // Else Send 0
  pause; // only construct to consume logical time
 }
}
>< // asynchronous operator
{ // Clock-domain 2
  while(true){ receive S; /* do something with S */}
}
\end{lstlisting}
\end{SubFloat}
\begin{SubFloat}{\label{delaycode:b}Timing diagram -- the X-axis shows
    the execution of the \texttt{wait\_inbetween} statement. The Y-axis
    is the time elapsed.}
  % \begin{figure}[h!]
    \centering
    \includegraphics[scale=0.7]{FF1}
  % \end{figure}
\end{SubFloat}
\caption{Programming a human responsiveness system (HRTCS) with
  \texttt{wait\_inbetween}}
% }
% \end{center}
\label{fig:1}
\end{figure}

Consider a machine used to test human response time and collect this
data for future research purposes. The machine switches on a green light
on a touch screen for anywhere between 50.3 ms to 200.3 ms, if the user
can touch this green light then she is successful and an integer
\texttt{1} is sent to a database, if unsuccessful a \texttt{0} is sent
instead. Different statistics and analyses on human responsiveness can
be based on the collected data.

The SystemJ pseudo-code~\cite{amal10} for such a system, along with the
expected timing behavior of the \texttt{wait\_inbetween} statement, is
shown in Figure~\ref{fig:1}. We use the SystemJ language in this paper,
first of all, because being a \textit{Globally Asynchronous Locally
  Synchronous} (GALS) language it helps us easily describe a larger
class of systems (like HRTCS in Figure~\ref{fig:1}) compared to purely
synchronous languages like Esterel. Moreover, being a super-set of
Esterel, SystemJ allows us to explore the marriage of real-time and
logical time, which is equally applicable to the synchronous sub-set.

The SystemJ program is divided into two clock-domains, the first
clock-domain continuously displays a green light (\texttt{GREEN\_LIGHT}
signal) on the touch screen sensor and waits for the user to respond. If
the user is able to touch the screen within the specified time of 50.3
ms to 200.3 ms a positive response is sent to the second
clock-domain. All the SystemJ programming constructs used to implement
this system are described in~\cite{amal10} and Table~\ref{tab:1}. The
\texttt{wait\_inbetween} specification, which models the non exact delay
between 50.3 ms and 200.3 ms is not supported in the current language
and cannot be compiled with the current SystemJ compiler. We introduce
real-time \texttt{wait\_inbetween} statements in SystemJ. Such non exact
real-time mechanisms are useful in developing systems where the
real-time delay is not know or should not be known a priori.

\subsubsection{Programming a timeout using the non exact real-time
  construct the \texttt{wait\_atleast} statement}
\label{sec:progr-time-using}

\begin{figure}[b!]
	\centering
	\vspace{-10pt}
        \begin{SubFloat}{\label{dd:a}Timeout waiting for the input
            signal \texttt{DoorOpened}}
        \begin{lstlisting}[style=sysj,morekeywords={abort,await,emit,present,trap,pause,exit,wait_atleast,suspend}]
trap(T){
  abort(DoorOpened){
    wait_atleast(10000 ms);
    exit(T);  
  }
}
\end{lstlisting}
\end{SubFloat}
\begin{SubFloat}{\label{dd:b}Timing diagram -- the X-axis shows the
    execution of the \texttt{wait\_atleast} statement. The Y-axis is the
    time elapsed.}
\includegraphics[scale=0.7]{FF2}
\end{SubFloat}
\caption{Programming a timeout with \texttt{wait\_atleast}}
\label{dd}
\end{figure}

There are many instances when one would like to wait on an input from
environment for only a specified amount of time. If the signal is not
received then a timeout is generated so that the program can make
progress. Such a timeout programmed in the SystemJ pseudo-code along
with its timing behavior is shown in Figure~\ref{dd}. The SystemJ
program is waiting on an input signal, \texttt{DoorOpened}, from the
environment for 10 seconds and makes progress irrespective of the
reception of this signal after 10 seconds.

The timing diagram, Figure~\ref{dd:b}, shows that the
\texttt{wait\_atleast} statement waits for a \textit{at least} 10
seconds, but unlike \texttt{wait\_inbetween} there is no upper
bound. The statement following the \texttt{wait\_atleast} statement is
executed after a finite countable, but unbounded delay. Such a statement
is useful for programming low priority tasks.

\subsubsection{Programming periodic tasks using the exact real-time
  construct the \texttt{wait\_exactly} statement}
\label{sec:progr-using-exact}

The previous wait statements are non exact, i.e., the execution of the
statement following the wait statement cannot be controlled exactly. The
final variant we introduce allows doing exactly that. We call it
\texttt{wait\_exactly} and this statement can be used to program exact
timeouts, periodic tasks, etc.

\begin{figure}[t!]
  \centering
	\vspace{-10pt}
        \begin{SubFloat}{\label{pp:a}Periodically emitting signal S}
        \begin{lstlisting}[style=sysj,morekeywords={emit,trap,pause,exit,wait_exactly}]
while(true) { 
 wait_exactly (1 ms); 
 emit S; 
 //do something 
}
\end{lstlisting}
\end{SubFloat}
\begin{SubFloat}{\label{pp:b}Timing diagram -- the X-axis shows the
    execution of the \texttt{wait\_exactly} statement. The Y-axis is the
    time elapsed.}
  \includegraphics[scale=0.7]{FF3}
\end{SubFloat}
  \caption{Programming a periodic task with wait\_exactly}
  \label{fig:periodic}
  \vspace{-10pt}
\end{figure}

One such example; periodic emission of a signal to the environment is
shown in Figure~\ref{fig:periodic}. As before upon execution of
\texttt{wait\_exactly (1 ms)} statement, the time starts elapsing. But,
unlike in previous cases the following statement needs to be executed
once 1 ms have passed, as shown in Figure~\ref{pp:b}.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: