### 2014-01-27 related_work.tex

\section{Related work}
\label{sec:related-work}

A number of studies have been performed when it comes to integrating
real-time directly as a programming construct in GALS (and its subset
synchronous) reactive languages. The most prominent works in this area
are by Shyamsundar~\cite{rsh94} and Bourke et
al.~\cite{Bourke2009a}. Shyamsundar incorporates real-time using
external timers in \textit{Communicating Reactive Processes} (CRP),
which like SystemJ is an extension of synchronous language Esterel to
asynchorny. But, as mentioned in Section~\ref{sec:inter-timers-react}
and~\ref{sec:resolution-real-time} external timers do not interact well
with preemption constructs in these languages. Bourke et al. introduce
real-time as first class constructs in the Esterel language. They like
us provide real-time wait statements called \texttt{delay} as first
class Esterel programming constructs, and translate them into Esterel
kernel constructs. But, unlike us they do not translate wait statements
into \texttt{pause} constructs directly. Instead, logical ticks are
generated by using abstract notion of \texttt{event} and \texttt{sample}
platform dependent timers. This notion makes it a complex and inflexible
solution since the number of platform timers with certain resolutions
need to be determined and present on the system for the solution to be
realizable. Moreover, non exact real-time waits and integration with
non-maximal parallelism is not studied at all. Our solution does not
require external timers (like~\cite{rsh94}) or logical notion of timers
(like~\cite{Bourke2009a}). In our solution we calculate the real-time
waiting time in logical ticks \textit{after} allocation and scheduling,
it gives the compiler developers the chance to optimize for many
different criteria such as computation time, energy, power, etc, without
worrying about or violating the real-time wait specification.

Other works such as Quartz~\cite{glog02} and Bertin et
al.~\cite{Bertin:2000:TVR:1947412.1947439}, also incorporate real-time
wait statements in Esterel using \texttt{pause} and special
\textit{pragmas} (annotations), respectively.  The Quartz solution is
targeted at studying timing properties using model checkers rather than
implementation, like us. The solution introduced by Bertin et
al.~\cite{Bertin:2000:TVR:1947412.1947439} translates real-time pragmas
into Timed automata~\cite{alur94} for model-checking real-time
deadlines. Our solution on the other hand removes all continuous time
elements and produces single discrete time model thereby making the
solution amenable to existing functional and real-time verification
tools developed for reactive languages and more in the spirit of logical
time. Finally, a major difference between Bertin et al. and our approach
is that their approach requires programmers to annotate the time taken
by `C' code fragments manually, this is extremely error prone, we use
static low-level program analysis to find out such times automatically.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: