### 2014-01-28 delays.tex
updated text

\section{Introducing real-time in SystemJ}
\label{sec:intr-real-time}

We introduce a single \textit{derived} statement called
\mbox{\texttt{wait\_inbetween (M..N)}}, built from kernel constructs, in
the SystemJ language for real-time control. The other two constructs
\mbox{\texttt{wait\_atleast (M)}} and \mbox{\texttt{wait\_exact (M)}}
are variants of this single derived construct.

% The resolution of the delay statement is of secondary concern and is
% dependent upon the execution platform.

\subsection{Semantics of the \texttt{wait\_inbetween} statement}
\label{sec:semant-delay-stat}

% \begin{figure}[t!]
%   \centering
%   \includegraphics[scale=0.6]{./sem}
%   \caption{Pictorial representation of the semantics of the delay
%     statement and its variants}
%   \label{fig:5}
% \end{figure}

Given a SystemJ program: \texttt{wait\_inbetween (M..N); p}, where $M
\in \mathbb{Q}^{>0}$ and $N \in \mathbb{Q}^{>0}$, statement $p$ is
executed after real-time delay of $\tau$ units, such that, $M \leq \tau
\leq N$.

We also introduce two variants of the derived statement
\texttt{wait\_inbetween}.
\begin{enumerate*}
\item Given a SystemJ program \texttt{wait\_atleast (M); p}, where $M
  \in \mathbb{Q}^{>0}$, statement $p$ is executed after real-time delay
  of $\tau$ units, such that, $M \leq \tau < \infty$. It is important to
  note that the lower bound of the wait construct; $M$ is \textit{not}
  an exact delay, but rather the control is allowed to proceed to the
  next statement anytime after waiting for \textit{at least} $M$ times
  units.
\item Given a SystemJ program \texttt{wait\_exact (M); p}, where $M
  \in \mathbb{Q}^{>0}$, statement $p$ is executed after real-time delay
  of $\tau$ units, such that, $M \leq \tau \leq M$. In this variant, the
  waiting time is exact.
\end{enumerate*}

% Figure~\ref{fig:5} pictorially represents the semantics of these delay
% constructs. The construct \texttt{delay (M..N)} requires that upon
% execution, the control-flow is suspended for \textit{at least} $M$ units
% and \textit{at most} $N$ units, depicted as bounded box in
% Figure~\ref{fig:5}. There is no upper bound for the delay variant
% \texttt{delay(M)}, which requires that the control-flow suspends for
% \textit{at least} $M$ units, but the following statement can be executed
% after some countable delay, depicted with the unbounded box in
% Figure~\ref{fig:5}. Finally, \texttt{delay (M..M)} requires that there
% is an \textit{exact} delay of $M$ units, i.e., the control-flow suspends
% execution for \textit{at least} and \textit{at most} $M$ units.

\subsection{Rewriting the \texttt{wait\_inbetween} statement}
\label{sec:rewr-delay-stat}

The introduced \texttt{wait\_inbetween} construct is not a kernel
construct, but a \textit{derived} construct built from the kernel
constructs (Table~\ref{tab:1}) in SystemJ. Figure~\ref{fig:3} gives the
rewrite of the \texttt{wait\_inbetween} construct to kernel statements.

\begin{figure}[tb]
	\vspace{-10pt}
    \begin{minipage}{\textwidth}
      \begin{lstlisting}[style=sysj,morekeywords={abort,await,emit,present,trap,pause,exit,delay,suspend}]
trap(T){
 int x = 0;
 while(true){
  x = x + 1;
  pause;
  if(x == d) exit (T); //wait for "d" ticks
 } // d is the number of logical ticks calculated
} // by Algorithm 1
\end{lstlisting}
    \end{minipage}
    \caption{The rewrite of \texttt{wait\_inbetween} construct}
    \label{fig:3}
\end{figure}

The fundamental observation is that real-time is converted into logical
time via the \texttt{pause} construct. The rewrite basically maps the
physical notion of time back to the elegant logical notion of time. The
rewrite \textit{waits} a certain number of logical ticks, before
proceeding to the next statement. The number of logical ticks \texttt{d}
to \textit{wait} (Figure~\ref{fig:3}) is determined by the compiler
statically at compile time. The value of \texttt{d} is intricately tied
with the WCRT and BCRT of the program and hence the execution platform.

\subsection{Finding the logical wait \texttt{d}}
\label{sec:find-logic-delay}

\begin{algorithm}[t!]
  \begin{minipage}{1.0\linewidth}
    \SetAlgoLined
    \KwData{WCRT, BCRT, $M \in \mathbb{Q}^{>0}$, $N \in \mathbb{Q}^{>0}$}
    \KwResult{d}
    let $l_1 \leftarrow \lceil \frac{M}{WCRT} \rceil$\;
    let $l_2 \leftarrow \lceil \frac{M}{BCRT} \rceil$\;
    let $u_1 \leftarrow \lfloor \frac{N}{WCRT} \rfloor$\;
    let $u_2 \leftarrow \lfloor \frac{N}{BCRT} \rfloor$\;
    let $F:(l_1,u_1) \rightarrow S_1$\;
    let $F:(l_2,u_2) \rightarrow S_2$\;
    let $D \leftarrow S_1 \cap S_2$\;
    \Return (some $d \in D$)\;
    \caption{Finding the value of \texttt{d}}
    \label{alg:1}
  \end{minipage}
\end{algorithm}

The computation of \texttt{d} is shown in Algorithm~\ref{alg:1}. This
algorithm is carried out for each CD (in SystemJ) or a synchronous
program individually. \textit{The fundamental observation is -- the
  reaction time for each logical tick is elastic -- varying only between
  the BCRT and the WCRT, thus any number of logical ticks \texttt{d}
  that map to the required real-time \mbox{\texttt{wait\_inbetween
      (M..N)}} should be chosen in such a way that they are invariant to
  this elasticity.}

Our algorithm takes as input: WCRT, BCRT, and the lower and upper bounds
$M$ and $N$, respectively, of the \texttt{wait\_inbetween} construct. We
first divide $M$ and $N$ with the BCRT and WCRT, respectively. This
division gives us the number of individual logical ticks required to
postpone the CD (or synchronous program) by the real-time
specification. We always $ceil$ when dividing $M$ and $floor$ when
dividing $N$ to make sure that the resultant values are integers (in
domain $\mathbb{N}^{>0}$) and these functions guarantee that the
resultant logical ticks result in real-time postponements between the
required range $[M,N]$. Next, a function $F$ maps these calculated
values to a set of equidistant integer points (values) separated by a
unit value -- these points represent all the logical ticks running at
the WCRT and the BCRT, respectively that satisfy the real-time wait
requirements. The intersection of these two sets gives all the logical
ticks that satisfy the real-time requirements invariant of the logical
time and its elasticity.

Let us revisit the HRTCS example (cf. Figure~\ref{fig:1}) to illustrate
the algorithm. From Figure~\ref{fig:1} we know that $M$ is 50.3 ms and
$N$ is 200.3 ms, respectively. \hj{Timing analysis of the program has
found} WCRT and BCRT be: 0.112 ms and 0.0334 ms, respectively. Thus, the
algorithm proceeds as follows:

\begin{enumerate*}
\item $l_1 \leftarrow \lceil 50.3/0.112 \rceil$ and $u_1 \leftarrow
  \lfloor 200.3/0.112 \rfloor$. $l_1 = 450$ and $u_1 = 1788$. We first
  calculate the logical ticks that are always running at the WCRT and
  satisfy the required real-time wait period.
\item $l_2 \leftarrow \lceil 50.3/0.0334 \rceil$ and $u_2 \leftarrow
  \lfloor 200.3/0.0334 \rfloor$. $l_2 = 1506$ and $u_2 = 5997$. We do
  the same for the BCRT case.
\item $S_1 = [450,1788]$ and $S_2 =[1506,5997]$. We then map the
  resultant bounds to linear points. Sets $S_1$ and $S_2$ represent
  logical ticks that running at the WCRT and BCRT, respectively always
  satisfy the required real-time constraints.
\item $D = S_1 \cap S_2$, $D = [1506,1788]$. Finally, the intersection
  of the two sets gives the set $D$ from which we can choose any value
  for \texttt{d}.
\end{enumerate*}

\begin{figure}[t!]
  \centering
  \includegraphics[scale=0.6]{FF4}
  \caption{The pictorial representation of the solution for the HRTCS system}
  \label{fig:hrtcssoln}
\end{figure}

The resultant value for \texttt{d} gives the number of logical ticks,
which can run at any physical clock-speed, bounded by the BCRT and the
WCRT of the program and still result in the desired real-time wait
period. Figure~\ref{fig:hrtcssoln} gives the solution space for the
aforementioned calculation. The patterned space is the only overlapping
region and hence gives the solution space for the above example.

We think this is an elegant solution, because the technique provides
real-time guarantees while preserving the essence (elastic logical tick)
of synchronous and GALS programming prescribed by SystemJ and Esterel
style languages. Moreover, this technique considers non
maximal-parallelism, i.e., the postponement in logical ticks is
calculated after scheduling of CDs or synchronous programs has been
performed. To our knowledge we are the first to do so.

\subsection{Extending the technique to variants of wait}
\label{sec:extend-tehcn-vari}

\paragraph{The \texttt{wait\_atleast(M)} construct}
\label{sec:extend-techn-vari}

The \texttt{wait\_atleast(M)} variant is easily accommodated in the
technique. All one needs to do is find the set $S_1$ and choose a value
from this set.

\paragraph{The \texttt{wait\_exact(M)} construct}
\label{sec:extend-techn-vari}

The \texttt{wait\_exact(M)} variant is more interesting. Like before,
we find sets $S_1$ and $S_2$, and find the intersection of the two sets
to get the value of \texttt{d}. It is possible % (and often likely, as
% suggested by our experiments in
% Section~\ref{sec:experimental-results})
that the resultant set $D$ is empty (also possible in the case of
\texttt{wait\_inbetween(M..N)}, but never possible in case of
\texttt{wait\_atleast(M)}). Note that $S_1$ and $S_2$ can never be empty
sets and will always contain at minimum a single element. For such cases
we provide two solutions listed below.

% In such a case, we can \textit{relax} the upper bound of the
% \texttt{delay} statement.

\paragraph{Relaxation of the upper real-time bound}
\label{sec:over-appr-relax}

The relaxation technique is shown in Algorithm~\ref{alg:2}.

\begin{algorithm}[t!]
  \begin{minipage}{1.0\linewidth}
    \SetAlgoLined
    \KwData{$S_2$, $D$, WCRT}
    \KwResult{d}
    \If {$D = \emptyset$} {
      let $j_{0}$ be the first element of set $S_2$\;
      \ShowLn let $N \leftarrow WCRT \times j_0$\;
      let $d \leftarrow j_0$\;
    }
    \Return d\;
    \caption{Calculating the minimum relaxation of the upper real-time
      bound}
    \label{alg:2}
  \end{minipage}
\end{algorithm}

This algorithm results in the \underline{smallest} relaxation required
for the real-time postponement to be satisfied. The algorithm takes as
input the WCRT and set $S_2$, recall that set $S_2$ represents the
logical ticks required to satisfy the real-time requirement at the
BCRT. We take the very first value from set $S_2$ and multiply it with
the WCRT to get the relaxation $N$. The first element of set $S_2$ is
returned as the logical tick \texttt{d}. The fundamental observation is
that we have to postpone the proceeding statement for at least $M$ units
of real-time, hence, under-approximation is out of question. We can
still over-approximate, but to reduce the resultant error, we should
over-approximate by the least possible value, which is the lower bound
of set $S_2$. Thus, the lower bound is considered to be the only element
shared between the two sets $S_1$ and $S_2$ and accordingly, the upper
real-time bound is relaxed by the multiplication of WCRT and the first
element of set $S_2$.

For example, assume $BCRT=5$ and $WCRT=100$, with a required
postponement of $(200..200)$. Algorithm~\ref{alg:1}, results in $D$
being an empty set, since $S_1=[2]$ and $S_2=[40]$, respectively. In
such a case, the resultant relaxation is: $40 \times 100$ using
Algorithm~\ref{alg:2}, which results in a time of $4000$ units. Such
large over-approximations can be avoided by breaking the critical paths
(worst case ticks) in the program (by inserting \texttt{pause}
constructs), just like in hardware design, which is very
\textit{unsurprising} considering Esterel has its origins in hardware.

\paragraph{Periodic execution of the clock domains}
\label{sec:peri-exec-clock}

The aforementioned technique maintains the essence of synchronous and
GALS programs -- the logical tick, but in doing so is unable to provide
the exact real-time wait of $M$ time units in case of
\texttt{wait\_exact(M)} statement, thereby violating its
semantics. The second option is to execute the CD periodically at the
WCRT. WCRT determines the fastest execution speed of any CD with the
guarantee that no input events from the environment will be missed
(cf. Section~\ref{sec:mapping-logical-time}). Thus, if we simply execute
the CD at periodically at the WCRT, we can choose the value of
\textit{d} from set $S1$. This technique allows meeting the synchronous
hypothesis, while at the same time guaranteeing the semantics of the
\texttt{wait\_exact(M)} statement. But, compared to the previous
option, this technique is very inflexible and results in reduced
response time and increased power consumption.

We believe that the designer should be allowed to choose from either of
these techniques and hence, provide this option via compiler switches.

\subsection{The tool-chain flow}
\label{sec:tool-chain-flow}

\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.45]{./tool.pdf}
  \caption{The tool-chain flow of SystemJ compiler with waits}
  \label{fig:4}
\end{figure}

The interaction of the low-level execution platform and the described
high-level technique of introducing real-time waits in the SystemJ
language is shown in Figure~\ref{fig:4}. We use static-analysis
techniques to find out the WCRT/BCRT of the programs automatically after
rewrites and standard compiler optimizations. We require timing
analyzable platforms, in our case \textit{Java Optimized Processor}
(JOP)~\cite{jop:jnl:jsa2007}. Moreover, since algorithms~\ref{alg:1}
and~\ref{alg:2} are performed after scheduling for different constraints
such as faster execution time, or reduced power consumption, we,
inherently satisfy the criteria of exploiting non-maximal
parallelism. The low-level analysis guarantees a precise WCRT/BCRT for
the logical tick, which is then used for calculating $d$ in
Figure~\ref{fig:3}.

Our compiler has a switch that enables over-approximation as described
in Section~\ref{sec:extend-tehcn-vari} provided the wait requirements
are not met. If this compiler switch is not enabled, the compiler simply
decides to execute the CD periodically, where the period is the
WCRT. WCRT effects the over-approximation or the period of the CD,
irrespective of the solution that the system designer chooses. A large
WCRT results in reduced performance and hence it is essential that the
WCRT be optimized. One can optimize the WCRT of any CD by breaking the
critical paths using the \texttt{pause} construct as we show later in
Section~\ref{sec:experimental-results}.

% The system designer can then either set the compiler switch or change
% the program by shortening the critical path by introducing
% \texttt{pause} constructs. We show how introducing \texttt{pause}
% constructs can help reduce the WCRT in
% Section~\ref{sec:experimental-results}.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: