### 2014-01-28 introduction.tex

\section{Introduction and motivation}
\label{sec:intr-motiv}

Reactive languages~\cite{gber931,amal10} are a class of programming
languages used for designing and implementing reactive systems, which
continuously respond to input from their environment. These languages
have been successfully used in programming a plethora of systems such as
fly-by-wire in Airbus~\cite{eairbus}, security surveillance
systems~\cite{amal121}, etc. Usually these systems also need to meet
real-time constraints imposed by the environment. Yet, these languages
do not support describing real-time statements as first class language
constructs.  For example, one cannot describe a simple real-time delay
(postponement) of an operation for 0.2 ms. These languages are based on
formal semantics, essential for the formal reasoning about \hj{code
generation} and verification of correctness of functional properties of
the developed programs, but leave nonfunctional properties such as
timing behavior as an implementation detail~\cite{boldt07}. Arguably,
rightly so, because physical time cannot be incorporated without
information about the underlying execution platform.  But, one can still
reason in terms of logical time. These languages support discrete
logical clock rather than a discrete physical (real-time) clock. The
real-time period of the discrete logical clock (also referred to as a
logical tick) is not fixed and it is determined by the responsiveness of
the program to the input signals. Unlike a discrete physical clock,
which has fixed real-time period, the period of the logical clock is
elastic. Although the timing model with logical clock has worked well
for reactive synchronous/GALS languages in designing discrete systems,
there is a need to introduce real-time since often, implementation
models are in the continuous or discrete real-time
domain~\cite{DBLP:journals/pieee/SifakisTY03}.  Moreover, adding timing
capabilities as first class language constructs create additional
opportunity for formally verifying functional and non-functional
requirements before system deployment.

\input{motivation}

The major motivations for introducing these real-time \textit{wait}
mechanisms and at the same time \textbf{contributions} of the paper are:
\hj{(1) real-time wait does not affect the model of logical time in
	SystemJ program as long as the amount of delay is within boundaries
	that can be determined statically by program analysis,} (2) the
	waiting model relies on the use of relative instead of absolute
	real-time, i.e., a wait in selected time units is counted from the
	currently executing SystemJ statement and (3) these waiting mechanisms
	allow mixing behaviors with real-time features and others with only
	logical time. All these as well as semantics of the wait constructs
	are discussed and illustrated in the rest of the paper. Our
	contributions can be further refined as follows:

\begin{enumerate*}
\item Real-time is modeled in the $\mathbb{Q}^{>0}$ domain of numbers,
 i.e., non-negative rational numbers.
\item We consider the property of \textit{non
    maximal-parallelism}. \textit{Non-maximal parallelism:} There are
  not always sufficient resources for processes to execute, so
  effectively processes are allocated and scheduled with different
  optimization criteria -- reducing computation time, reducing power
  consumption, etc. This allocation and scheduling is an integral part
  of any language compilation and needs to be considered when
  introducing real-time waits.
\item We are the first, to our knowledge, to allow specification of
  both; exact and non-exact real-time in GALS and synchronous programs.
\item We \hj{clarify} upon a key insight relating real-time and
  logical-time -- \textit{the resolution of any real-time statement is
    dependent on the worst case logical tick time} -- an aspect that
  seems to have escaped in the current literature on incorporation of
  real-time in synchronous/GALS languages.
\end{enumerate*}

The rest of the paper is arranged as follows:
Section~\ref{sec:background} gives the background on SystemJ and related
techniques required for the rest of the
paper. Section~\ref{sec:intr-real-time} introduces real-time mechanisms
in the GALS paradigm and describes their compilation into logical
time. Section~\ref{sec:proof-correctness} sketches the correctness
proofs for the introduced
algorithms. Section~\ref{sec:experimental-results} gives the
experimental results for a set of real-time applications. In
Section~\ref{sec:disc-perc-limit} we discuss the advantages and
limitations of the proposed approach. Section~\ref{sec:related-work}
positions this work in relation to other approaches to introducing
real-time. Finally, we conclude in Section~\ref{sec:concl-future-work}.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: